---
title: "Netfilx Analysis"
author: "Yujia Dai"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(ggplot2)
```

## Description
The [dataset](https://www.kaggle.com/datasets/shivamb/netflix-shows) contains movies and TV shows that released on the Netfilx. The dataset has 12 columns and 8807 rows. 

## Reading data

```{r read dataset}
#setwd("/Users/daidai/Desktop/ds_project/netflix")
df <- read.csv("netflix_titles.csv")
head(df, n = 5)

# check dataset size and structure
dim(df)
str(df)
```

The dataset contains 8807 rows and 12 columns. The table shows there are some values is missing. From $str(df)$, we notice that the empty values is shown as "" rather than na, so $is.na()$ doesn't work. 

## Data cleaning

```{r data read}
# colMeans() function in R calculates the mean for each column of a matrix or a dataframe
colMeans(df == "") * 100
sum(duplicated(df))
```


```{r eda1}
# get first 10 rows with empty values in director
missing_rows <- df[df$director == "", ]
print(head(missing_rows), 10)
```

```{r eda2}
# fill empty values in 'director', 'cast' and 'country' features with 'unknown'
df$director <- ifelse(df$director == "", "Unknown", df$director)
print(head(df$director), 10)

df$cast <- ifelse(df$cast == "", "Unknown", df$cast)
print(head(df$cast), 10)

df$country <- ifelse(df$country == "", "Unknown", df$country)
print(head(df$country), 10)
```

```{r}
# drop empty value in columns
df <- df[apply(df, 1, function(row) !any(row == "")), ]
colMeans(df == "") * 100
sum(duplicated(df))
```

### Split Added Date
```{r}
# split added date and added year in 'date_added'
date <- strsplit(df$date_added, ", ")
added_date <- sapply(date, function(x) x[1])
added_year <- sapply(date, function(x) x[2])
df$added_month <- added_date
df$added_year <- added_year

# remove 'date_added' column
df <- df[, !(names(df) %in% c("date_added"))]
```

```{r}
# drop rows that 'added_year' is less than 'release_year'
df <- df[!(df$added_year < df$release_year), ]
dim(df)
```

## Analysis features
### Video Type
```{r}
type_counts <- table(df$type)
type_counts
ggplot(df, aes(x = type, fill = type)) +
  geom_bar(width = 0.5) +
  labs(x = "Type", y = "Count") +
  theme_minimal()
```

```{r}
movie_percentage <- round((type_counts["Movie"] / sum(type_counts)) * 100, 2)
tv_show_percentage <- round((type_counts["TV Show"] / sum(type_counts)) * 100, 2)

sprintf("On Netflix, %.2f%% are Movie and %.2f%% are TV show", movie_percentage, tv_show_percentage)
```

### Director
```{r}
director_counts <- sort(table(df$director), decreasing = TRUE)
director_counts[1:10]
```

### Country
```{r}
country_counts <- sort(table(df$country), decreasing = TRUE)
top_countries <- as.data.frame(country_counts[1:5])

ggplot(top_countries, aes(x = Var1, y = Freq, fill = Var1)) +
  geom_bar(stat = "identity") +
  labs(x = "Country", y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Date
```{r}
date_counts <- sort(table(df$added_year))
date_df <- data.frame(Year = as.integer(names(date_counts)), Count = as.integer(date_counts))
date_df$Year <- as.factor(date_df$Year)


ggplot(date_df, aes(x = Year, y = Count, fill = Year)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
### Rating
```{r}
sort(prop.table(table(df$rating)), decreasing = TRUE)
ggplot(df, aes(x = rating, fill = rating)) +
  geom_bar() +
  coord_flip() +
  labs(y = "Count", x = "Rating") +
  theme_minimal()
```

### Top five countries preference for video's type
```{r}
top_five <- top_countries$Var1
df<-df[df$country %in% top_five, ]
df$country <- factor(df$country, levels = top_five)

ggplot(df, aes(x = country, fill = type)) +
  geom_bar(position = "dodge") +
  labs(x = "Country", y = "Count", fill = "Type") +
  theme_minimal()
```

